---
title: "Bispecific Analysis"
author: "Bryan"
date: "2022-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, message = F, warning = F}

library(tidyverse)
library(kableExtra)
library(cowplot)
library(viridis)
library(VISCPK)

theme_set(theme_bw() + theme(legend.position = "top", legend.key.width = unit(1, "cm"),
                             legend.box = "vertical"))
source("code/neutralization-funs.R")
source("code/onecmptPKPD-sims.R")

ratio_lab = "mAb1/total dose"
ratio_breaks = 0:4/4
ratio_labels = c("mAb 2", 1:3/4, "mAb 1")
ratio_labels_x = c("All\nmAb 2", 1:3/4, "All\nmAb 1")

doses = c(150, 300, 600, 1200, 2400)

interaction_labeller = function(interaction_var){
  factor(interaction_var,
         labels = c("Min. Neut.", "Max. Neut.", "Additivity", "Bliss-Hill"),
         levels = c("minNeut", "maxNeut", "additivity", "BH")
  )
}

# parameter grid

hl = c(7, 28, 42, 84)
omega = c(0.33, 0.67, 1)
muIC50 = c(-3, -2, -1)
sigIC50 = c(0.25, 0.5, 1)

full_combinations = crossing(
  hl_1 = hl,
  hl_2 = hl,
  omega_1 = omega,
  omega_2 = omega,
  muIC50_1 = muIC50,
  muIC50_2 = muIC50,
  sigIC50_1 = sigIC50,
  sigIC50_2 = sigIC50
) %>%
  filter(hl_1 <= hl_2)


main_vars = c("ratio", "c0", "muIC50_1", "muIC50_2",
               "hl_1", "hl_2", "sigIC50_1", "sigIC50_2",
              "omega_1", "omega_2")
```


```{r load-data}
load("output/bispecific_data.rda")
```

## Setup

Combine a superb bnab with bad half-live with decent mab with great half-life.

- mab1: muIC50 = -3, sd (log10 IC50) = 0.25, resistance = 0%, hl = 7
- mab2: muIC50 = -2, sd (log10 IC50) = 0.5, resistance = 33%, hl = 84

Here we will want to look at both mean IIP and coverage IIP > 2

```{r CS2-setup}
#the case study2 notation is an artifact of old two case study analysis

case_study2_full = bispecific_data %>%
  filter(muIC50_1 == -3 & muIC50_2 == -2 &
           sigIC50_1 == 0.25 & sigIC50_2 == 0.5 &
           omega_1 == 1 & omega_2 == 0.67 &
           hl_1 == 7 & (hl_2 == 7 | hl_2 == 84))

CS2_iip = case_study2_full %>%
  select(-contains("coverage")) %>%
  gather(key, value, -all_of(main_vars)) %>%
  separate(key, into = c("interaction", "endpoint", "summary", "time"), sep = "_") %>%
  unite(summary, summary, endpoint, sep = "_")


CS2_coverage = case_study2_full %>%
  select(all_of(main_vars), contains("coverage")) %>%
  gather(key, value, -all_of(main_vars)) %>%
  separate(key, 
           into = c("interaction", "endpoint", "coverage_endpoint", "summary", "time"), 
           sep = "_") %>%
  mutate(
    coverage_level = parse_number(coverage_endpoint),
    summary = "mean_coverage"
    ) %>%
  select(-endpoint, -coverage_endpoint)

# CS = case study
CS2_data = bind_rows(CS2_iip, CS2_coverage) %>%
  mutate(summary_label = if_else(
    summary == "mean_iip",
    "Mean IIP",
    paste("% viruses IIP >", coverage_level)),
    value = if_else(summary == "mean_iip", value, 100 * round(value, 2))
  )


#main_vars[-1] should exclude ratio
CS2_combo = CS2_data %>%
  filter(c0 != 2400, hl_1 != hl_2) %>%
  group_by_at(all_of(main_vars[-1])) %>%
  group_by(interaction, summary_label, summary, time, coverage_level, .add = T) %>%
  summarize(
    total_max = length(which(value==max(value))),
    best_ratio = ratio[which.max(value)],
    best_value = max(value),
    `.groups` = "drop"
  ) %>%
  mutate(
    admin = "combination",
    )
  
CS2_bispecific = CS2_data %>%
  filter(ratio == 0.5 & hl_1 == hl_2 & c0 > 150) %>%
  rename(
    best_ratio = ratio,
    best_value = value
  ) %>%
  mutate(
    admin = "bi-specific",
    c0 = c0/2
  )

CS2_summary = bind_rows(CS2_combo, CS2_bispecific) %>%
  ungroup() 
```

## Parental ratio optimization and PK

Look at all doses and summary endpoints to start. Switch to 300, 1200ug; mean iip

```{r dual-ratio, eval = F}

CS2_data %>%
  filter(c0 == 1200 & hl_1 != hl_2 & (is.na(coverage_level) | coverage_level == 2) &
           interaction == "BH") %>%
  dplyr::select(-coverage_level, -summary_label) %>%
  spread(key = summary, value = value) %>%
  ggplot(aes(x = ratio, y = mean_iip)) +
  geom_line() +
  geom_line(aes(y = mean_coverage * 6), colour = "blue") +
  scale_y_continuous(sec.axis = sec_axis(~./6, name="mean_breadth")) +
  facet_wrap( ~ time) +
  theme(axis.title.y.right = element_text(color = "blue", size=13),
        axis.text.y.right = element_text(color = "blue"))



```

```{r CS2-bestratio, fig.cap="supplement. Optimal ratios for all summaries and interactions with combination administration."}

CS2_summary %>%
  filter(admin == "combination") %>%
  ggplot(aes(x = interaction, y = best_ratio, shape = time)) +
  geom_point() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(breaks = ratio_breaks, labels = ratio_labels, limits = c(0, 1)) +
  labs(y = "optimal ratio") +
  facet_grid(rows = vars(summary_label), 
             cols = vars(c0), space = "free", scales="free_x") +
  theme(legend.box = "vertical", axis.text.x = element_text(angle = 30, hjust = 1))

```

```{r CS2-bestratio-subset, fig.cap="subset for analysis. not for paper."}

CS2_summary %>%
  filter(admin == "combination" & c0 %in% c(300, 1200) &
           (summary == "mean_iip" | coverage_level == 2)) %>%
  ggplot(aes(x = interaction, y = best_ratio, shape = time)) +
  geom_point() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(breaks = ratio_breaks, labels = ratio_labels, limits = c(0, 1)) +
  labs(y = "optimal ratio") +
  facet_grid(rows = vars(summary_label), 
             cols = vars(c0), space = "free", scales="free_x") +
  theme(legend.box = "vertical", axis.text.x = element_text(angle = 30, hjust = 1))


```

```{r ratio-sensitivity}

CS2_data %>%
  filter(hl_1 != hl_2 & c0 %in% c(300, 1200) &
           (summary == "mean_iip" | coverage_level == 2)) %>%
  ggplot(aes(x = ratio, y = value, colour = factor(c0), linetype = time)) +
  geom_line() +
  scale_x_continuous(breaks = ratio_breaks, labels = ratio_labels_x, limits = c(0, 1)) +
  geom_point(data = subset(CS2_combo, c0 %in% c(300, 1200) &
           (summary == "mean_iip" | coverage_level == 2)),
           aes(x = best_ratio, y = best_value)) +
  facet_grid(rows = vars(summary_label), 
             cols = vars(interaction),  scales="free_y", switch = "y") +
  labs(colour = "dose (mg)", linetype = "", x = "ratio mAb 1 to total dose", y= "") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        panel.spacing.x = unit(1, "lines"),
        strip.switch.pad.grid = unit(0, "lines"),
        legend.box = "horizontal")

```


```{r CS-HL, fig.cap = "what does the varying combination 1-cmpt PK look like. You basically get a fixzed mAb 2"}

CS2_summary %>%
  filter(c0 %in% c(300, 600) & summary_label %in% c("IIP > 2", "Mean IIP") &
           admin == "combination") %>%
  mutate(c0_1 = best_ratio * c0, c0_2 = (1-best_ratio) * c0) %>%
  group_by(c0, summary_label, interaction, time, best_ratio) %>%
  nest() %>%
  mutate(pk_data = map(data, ~.pk_1cmpt(select(., c0_1, c0_2, hl_1, hl_2), 84))) %>%
  filter(best_ratio > 0 & best_ratio < 1) %>%
  rename(timepoint = time) %>%
  unnest(cols = pk_data) %>%
  gather(mab, conc, conc_1, conc_2) %>%
  ggplot(aes(x = time, y = conc, colour = mab, linetype = factor(c0))) +
  geom_line() +
  scale_y_log10() +
  facet_wrap(~best_ratio)

```


```{r bispecific-pk-pl, fig.cap="bi-specific 1-cmpt PK"}
bispecific_pk = map_df(c(300, 1200), ~ tibble(
  time = seq(0, 84, by = 1),
  c0 = .,
  conc = . / 3 * exp(-time * log(2) / 7)
)) 

bispecific_pk %>% 
  ggplot(aes(x = time, y = conc,  linetype = factor(c0))) +
  geom_line() +
  scale_y_log10()


```

## No synergy bi-specific vs combination comparision

The PK penalty is evident in the 1:1 comparison, where the bi-specific would dominate by assumption for common PK.

```{r}
balance_comb = CS2_data %>%
  filter(c0 %in% c(300, 1200) &
           hl_1 != hl_2 & (is.na(coverage_level) | coverage_level == 2) &
           ratio == 0.5) %>%
  mutate(best_value = value, best_ratio = ratio,
         admin = "1:1 combination")

```

```{r raw-bispecific-performance-iip}

CS2_summary %>%
  bind_rows(balance_comb) %>%
  mutate(
    admin = if_else(admin == "combination", "optimal combination", admin)
  ) %>%
  filter(summary == "mean_iip" & c0 %in% c(300, 1200)) %>%
  dplyr::select(-total_max, -coverage_level) %>%
  ggplot(aes(x = interaction, y = best_value, colour = admin)) +
  geom_point(position = position_dodge(width = 0.5)) +
  #geom_abline() +
  labs(y = "mean iip", color = "") +
  facet_grid(cols = vars(time), rows = vars(c0), space = "free", scales="free_x")

```


```{r raw-bispecific-performance-coverage}

CS2_summary %>%
  bind_rows(balance_comb) %>%
  mutate(
    admin = if_else(admin == "combination", "optimal combination", admin)
  ) %>%
  filter(summary == "mean_coverage" & c0 %in% c(300, 1200), coverage_level == 2) %>%
  dplyr::select(-total_max) %>%
  ggplot(aes(x = interaction, y = best_value, colour = admin)) +
  geom_point(position = position_dodge(width = 0.5)) +
  #geom_abline() +
  labs(y = "% viruses IIP > 2", color = "") +
  facet_grid(cols = vars(time), rows = vars(c0), space = "free", scales="free_x")

```

## Synergy Analysis

### Background

#### Neutralization

Synergy is described as an "IC50" boost-factor ($\alpha > 1$), like an inverse-potency reduction. The neutralization calculations are done as follows using the titer definition, where i = parental arm:
$$\tau_i = \frac{C(t)}{IC50_i}$$

- Minimum: $$1 - [1+\alpha*min_i\{\tau_i\}]^{-1}$$

- Maximum: $$1 - [1+\alpha*max_i\{\tau_i\}]^{-1}$$

- Additivity:  $$1 - [1+\alpha\sum_i{\tau_i}]^{-1}$$

- BH:  $$1 - \prod_i[1+\alpha\tau_i]^{-1}$$

#### IIP

The conversions to IIP then follow nicely:

- Minimum: $$log_{10}[1+\alpha*min_i\{\tau_i\}]$$

- Maximum: $$log_{10}[1+\alpha*max_i\{\tau_i\}]$$

- Additivity:  $$log_{10}[1+\alpha\sum_i{\tau_i}]$$

- BH:  $$\sum_ilog_{10}[1+\alpha\tau_i]$$

### Generate synergy profile


```{r setup-data}

bispecific_pd = CS2_bispecific %>% 
  select(omega_1, omega_2, muIC50_1, muIC50_2, sigIC50_1, sigIC50_2) %>% 
  distinct()

```

```{r sim-synergy}

#interaction_function()

viral_iC50s_sims = .sample_IC50s(total_virus_sims = 500, pd_parms = bispecific_pd,
                              virus_seed = 822960625)



calc_bispecific_iip = function(conc, viral_ic50s, synergy_levels = c(1, 10, 100, 1000)) {
  map_df(synergy_levels, function(i) {
    viral_ic50s %>%
      mutate(
        titer_1 = conc / IC50_1,
        titer_2 = conc / IC50_2,
        min_titer = pmin(titer_1, titer_2),
        max_titer = pmax(titer_1, titer_2)
      ) %>%
      transmute(
        synergy_factor = i,
        maxNeut = log10(1 + max_titer * i),
        minNeut = log10(1 + min_titer * i),
        BH = log10(1 + i * titer_1) + log10(1 + i * titer_2),
        additivity = log10(1 + i * (titer_1 + titer_2))
      )
  }) %>%
    gather(interaction, iip, maxNeut, minNeut, BH, additivity) %>%
    mutate(coverage = iip > 2) %>%
          group_by(synergy_factor, interaction) %>%
          summarize_all(list(mean = mean))
}
  
bispecific_pkpd = bispecific_pk %>%
  group_by(c0, time) %>%
  mutate(ID50_dat = map(conc, calc_bispecific_iip, viral_ic50s = viral_iC50s_sims)) %>%
  unnest(cols = ID50_dat) %>%
  rename(mean_iip = iip_mean, mean_coverage = coverage_mean)

bispecific_pkpd_summary = bispecific_pkpd %>%
  arrange(c0, interaction, synergy_factor, time) %>%
  group_by(c0, interaction, synergy_factor) %>%
  summarize_at(vars(mean_iip, mean_coverage), list(
      auc = ~ pracma::trapz(time, .x) / max(time),
      trough = ~(.x[time == 84])
  )
  ) %>%
  gather(summary, best_value, mean_iip_auc:mean_coverage_trough) %>%
  separate(summary, into = c("tmp1", "tmp2", "time")) %>%
  unite(summary, tmp1, tmp2, sep = "_") %>%
  mutate(admin = "bi-specific", coverage_level = 2) %>%
  ungroup() %>%
  mutate(
    best_value = if_else(summary == "mean_iip", best_value, 100 * round(best_value, 2)),
    summary_label = if_else(
      summary == "mean_iip",
      "Mean IIP",
      paste("% viruses IIP >", coverage_level)
    )
  )

```

```{r bispecific-pkpd}


bispecific_pkpd %>%
  gather(outcome, value, mean_iip, mean_coverage) %>%
  ggplot(aes(x = time, y = value, colour = factor(synergy_factor))) +
  geom_line() +
  facet_grid(cols = vars(interaction), rows = vars(outcome, c0), scales = "free_y")

```


### Synergy comparison results

```{r synergy-iip, fig.width=7}

CS2_combo %>%
  mutate(synergy_factor = 1) %>%
  bind_rows(bispecific_pkpd_summary) %>%
  filter(summary == "mean_iip" & c0 %in% c(300, 1200)) %>%
  ggplot(aes(x = interaction, y = best_value, colour = factor(admin), shape = factor(synergy_factor))) +
  geom_point(aes(group = admin), position = position_dodge(width = 0.1), size = 2) +
  labs(y = "mean iip", color = "", shape = "synergy (fold IC50 boost)") +
  scale_x_discrete(limits = c("BH", "additivity", "maxNeut", "minNeut")) +
  facet_grid(cols = vars(time), rows = vars(c0), space = "free", scales="free_x")

```

```{r synergy-coverage, fig.width=7}

CS2_combo %>%
  mutate(synergy_factor = 1) %>%
  bind_rows(bispecific_pkpd_summary) %>%
  filter(summary == "mean_coverage" & c0 %in% c(300, 1200) & coverage_level == 2) %>%
  ggplot(aes(x = interaction, y = best_value, colour = factor(admin), 
             shape = factor(synergy_factor))) +
  geom_point(aes(group = admin), position = position_dodge(width = 0.1), size = 2) +
  labs(y = "% viruses IIP > 2", color = "", shape = "synergy (fold IC50 boost)") +
  scale_x_discrete(limits = c("BH", "additivity", "maxNeut", "minNeut")) +
  facet_grid(cols = vars(time), rows = vars(c0), space = "free", scales="free_x")

```

## Master figure

Combines the parental and synergy analysis.

```{r paper-plot, fig.height=7}

bilab_location = 1.45
final_ratio_breaks = c(0, 0.5, 1, bilab_location)
final_ratio_labels = c("All\nbNAb2", "1:1", "All\nbNAb1", "Bi-\nspecific")

opt_plot_results = CS2_combo %>%
  ungroup() %>%
  dplyr::filter(c0 %in% c(300, 1200) & (summary == "mean_iip" | coverage_level == 2) & 
                  interaction != "maxNeut") %>%
  mutate(
    ratio = best_ratio,
    value = best_value,
    interaction = fct_rev(interaction_labeller(interaction))
  )

bispecific_plot_res = bispecific_pkpd_summary %>%
  filter(synergy_factor < 1000 & interaction != "maxNeut") %>%
  mutate(
    ratio = bilab_location,
    value = best_value,
    interaction = fct_rev(interaction_labeller(interaction))
  )

make_comparison_plot = function(time_var){
  CS2_data %>%
    filter(
      hl_1 != hl_2 & c0 %in% c(300, 1200) &
        (summary == "mean_iip" | coverage_level == 2) &
        time == time_var & interaction != "maxNeut"
    ) %>%
    mutate(interaction = fct_rev(interaction_labeller(interaction))) %>%
    ggplot(aes(x = ratio, y = value, colour = factor(c0))) +
    geom_line() +
    scale_x_continuous(breaks = final_ratio_breaks, labels = final_ratio_labels,
                       minor_breaks = 0:4/4) +
    geom_point(data = dplyr::filter(opt_plot_results, time == time_var), 
               shape = 21, fill = "white") +
    geom_point(data = dplyr::filter(bispecific_plot_res, time == time_var),
                                    aes(shape = factor(synergy_factor)),
               position = position_dodge(width = 0.25)) +
    facet_grid(space = "free_x",
      rows = vars(summary_label),
      cols = vars(interaction),
      scales = "free",
      switch = "y"
    ) +
    labs(
      colour = "dose (mg)",
      linetype = "",
      x = "",
      y = time_var,
      shape = "synergy (fold IC50 boost)"
    ) +
    scale_color_viridis_d(option = "D", end = 0.75) +
    theme(
      strip.placement = "outside",
      strip.background = element_blank(),
      panel.spacing.x = unit(1, "lines"),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      strip.text.y = element_text(size = 10, margin = margin(0,0,0,0)),
      strip.text.x = element_text(margin = margin(0,0,0,0)),
      axis.text.x = element_text(size = 7),
      strip.switch.pad.grid = unit(0, "lines"),
      legend.box = "horizontal",
      plot.margin = unit(c(0, r = 1, 0, 0), "cm")
    )
}

auc_pl = make_comparison_plot("auc") + theme(legend.position = "none") 
trough_pl = make_comparison_plot("trough")

pl_legend = get_legend(trough_pl +
                         theme(legend.box = "vertical", legend.margin = margin(0, 0, 0, 0, "lines"),
                               legend.box.margin = margin(0, 0, 0, 0, "lines"),
                               legend.text = element_text(size = 8),
                               legend.spacing.y = unit(0, "lines"),
                               legend.title = element_text(size = 8)))

plot_grid(pl_legend,
          plot_grid(auc_pl, trough_pl + theme(legend.position = "none"), 
                                    nrow = 2),
          rel_heights = c(1.25, 12), nrow = 2)
          
```

```{r fig4-eps, fig.height=7, dev = "postscript", include = F}

plot_grid(pl_legend,
          plot_grid(auc_pl, trough_pl + theme(legend.position = "none"), 
                                    nrow = 2),
          rel_heights = c(1.25, 12), nrow = 2)

```
