---
title: "Ratio optimization for 3BNC117, 10-1074, theoretical-mutants, and VRC07-LS"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Overview

In this case study, we perform ratio optimization on three bnabs, 3BNC117, 10-1074, and  VRC07-523LS. We use theoretical LS variants of these antibodies by extending their half-live by 2-fold: 3BNC117-T and 10-1074-T. We optimize specifically using potency reduction dose-response relationships derived from Pegu and AMP studies. The protection is estimates for each bnab and combined using an independence model (ie, Bliss-Hill). For these theoretical trials, we consider a simple design of 600 mg administered subcutaneously with a 3 month follow-up window.

# PK for parental, -T variants, and VRC07-523LS

The PK for 3BNC117 and 10-1074 comes from [citation]. In brief, the PK parameters are derived from  two-compartment models fit to participants summarized as the mean across individuals. To derive the theoretical LS mutants (-T), we extend the half-life by 3-fold [need a citation, pettit?]. This was accomplished by taking one third of the parental clearance without modifying the distribution. All of the PK parameter estimates for VRC07-523LS were estimated in [lancet]. For SC design, distribution to the central compartment is delayed, distributing at rate of $ka$ post-injection. Bioavailablity may also be affected when SC administration is used. As this is an illustrative example and these parameters are not available in 3BNC117 or 10-1074, we use the VRC07-523LS estimates assuming that these properties are similar for anti-HIV mabs.


```{r, message = F, warning = F}

library(tidyverse)
library(kableExtra)
library(readxl)
library(cowplot)
library(viridis)
theme_set(theme_bw())

source("code/empirical-opt-funs.R")
source("code/neutralization-funs.R")

theme_set(theme_bw() + 
            theme(legend.position = "top", legend.key.width = unit(1, "cm"),
                  panel.grid.minor = element_blank()))

ic50_cutoff = 10
ic50_breaks = c(10^c(-3:1), ic50_cutoff * 10)
ic50_labels = c(0.001, 0.01, 0.1, 1, 10, "resistant")

parental_ratio_lab = "3BNC117:total dose"
parental_ratio_breaks = 0:4/4
parental_ratio_labels = c("Only\n10-1074", 1:3/4, "Only\n3BNC117")

ls_ratio_lab = "3BNC117-T:total dose"
ls_ratio_breaks = 0:4/4
ls_ratio_labels = c("Only\n10-1074-T", 1:3/4, "Only\n3BNC117-T")

ic50_title = expression(paste("IC"[50]," (", mu, "g/mL)"))

```

```{r opt-flags-setup}
# the optimizations are slow so don't want to run everytime
opt_2bnab_flag = !file.exists("output/double-opt.rda")  | !file.exists("output/double-opt-amp.rda")
opt_3bnab_flag = !file.exists("output/triple-opt.rda") | !file.exists("output/triple-opt-amp.rda")
```

```{r load-pk, message = F, warning = F}

parental_data = read_csv("data/empirical-pk-data/PK_parental_1074_3bnc.csv") %>%
  ungroup() %>%
  dplyr::select(analyte, parameter, value) %>%
  dplyr::filter(parameter %in% c("V1", "V2", "Cl", "Q")) %>%
  spread(parameter, value) 

tls_data = parental_data %>%
  mutate(
    Cl = Cl/3,
    analyte = paste0(analyte, "-T")
    ) 

vrc07_data = vrc_res <- read_table2("data/empirical-pk-data/PK_vrc07.txt") %>%
  janitor::clean_names() %>%
  mutate(analyte = "VRC07-523LS") %>%
  dplyr::select(analyte, parameter = variability, value = finalestimate)%>%
  spread(parameter, value) %>%
  rename(Cl = CL, ka = KA, V1 = V2, V2 = V3)

all_pk = parental_data %>%
  bind_rows(tls_data) %>%
  bind_rows(vrc07_data) %>%
  mutate(F1 = zoo::na.locf(F1), ka = zoo::na.locf(ka)) %>%
  rowwise() %>%
  mutate(
    rate_parms = list(convert_2cmp_pkparms(CL = Cl, Vc = V1, Vp = V2, Q = Q)),
    HL_elimination = log(2)/rate_parms$beta_parm
    ) %>%
  ungroup() 

all_pk %>%
  dplyr::select(-rate_parms) %>%
  rename(Vc = V1, Vp = V2, bioavailability = F1, `HL (days)` = HL_elimination) %>%
  gather(parameter, value, -analyte) %>%
  spread(analyte, value) %>%
  mutate(parameter = factor(parameter, 
                            levels = c("Vc", "Cl", "Q", "Vp", "ka", "bioavailability", "HL (days)"))
         ) %>%
  arrange(parameter) %>%
  select(parameter, `10-1074`, `3BNC117`, `10-1074-T`, `3BNC117-T`, `VRC07-523LS`) %>%
  kable(booktabs = T, digits = 2) %>%
  kable_styling(full_width = F) 

```

```{r example-pk, fig.cap = "Population PK for analytes with 600 mg administration."}

pk_window = seq(0, 84, by = 0.5)

example_pk_dat  = all_pk %>%
  unnest(cols = rate_parms) %>%
  mutate(dose = 600) %>%
  create_pk(pk_window) 

write_csv(example_pk_dat, "output/example-empirical-pk-dat.csv")

pk_pl_all = example_pk_dat %>%
  ggplot(aes(x = time, y = conc, color = analyte)) +
  geom_line() +
  labs(y = "Concentration (ug/mL; 600 mg dose)", x = "days")

pk_pl_all

```

```{r manuscript-pk}

man_pk_pl = all_pk %>%
  dplyr::filter(!analyte %in% c("10-1074", "3BNC117")) %>%
  unnest(cols = rate_parms) %>%
  mutate(dose = 600) %>%
  create_pk(pk_window) %>%
  ggplot(aes(x = time, y = conc, color = analyte)) +
  geom_line() +
  labs(y = expression(paste("concentration (",mu,"g/mL; 600 mg dose)")), colour = "", x = "days") +
  theme(legend.direction = "vertical", legend.position = c(1,1),
        legend.justification = c(1.05,1.05),      
        legend.key.height =  unit(0.5, "cm"),
        legend.margin = margin(t = -10),
         legend.box.background = element_rect(colour = "black"),
)

man_pk_pl

```

# PD

## CATNAP data

```{r catnap-profile}

# these inform Table 2: the sensitivity analysis parameters 

catnap_dat_raw = read_rds("output/processed_catnap_assay.rds") 

catnap_dat_raw %>%
  mutate(
    sensitive = clean_IC50 < ic50_cutoff,
    clean_IC50 = if_else(clean_IC50 >= ic50_cutoff, Inf, clean_IC50)
  ) %>%
  group_by(Antibody) %>%
  summarize(
    total_virus = n(),
    pct_resistant = mean(!sensitive),
    mean_sensitive = mean(log10(clean_IC50[sensitive])),
    sd_sensitive = sd(log10(clean_IC50[sensitive]))
  ) %>%
  filter(total_virus > 100) %>%
  gather(key, value, pct_resistant:sd_sensitive) %>%
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  facet_wrap(~key, scales = "free")
  

```


```{r pd-profile}

catnap_dat = catnap_dat_raw %>%
  subset(Antibody  %in% c("10-1074", "3BNC117", "VRC07-523-LS")) %>%
  group_by(Virus, Reference) %>%
  mutate(
    total_ab = n(),
    clean_IC50 = if_else(clean_IC50 >= ic50_cutoff, Inf, clean_IC50),
    clean_IC50_plot = if_else(clean_IC50 >= ic50_cutoff, ic50_cutoff * 10, clean_IC50)
  ) %>%
  filter(total_ab == 3) %>% # makes sure all three antibodies are tested (is this the best way to check this??)
  select(Virus, Antibody, clean_IC50, clean_IC50_plot, subtype, Reference) %>%
  ungroup()

# wide for pkpd analysis
pd_data = catnap_dat %>%
  dplyr::select(-clean_IC50_plot) %>%
  pivot_wider(names_from = "Antibody", values_from = "clean_IC50")


res_pl = catnap_dat %>%
  group_by(Antibody) %>%
  summarize(
    resistant = mean(clean_IC50 >= ic50_cutoff),
    resistant_frac = paste(n() * resistant, n(), sep = "/")
  ) %>% 
  ggplot(aes(x = Antibody, y = resistant * 100)) +
  geom_bar(aes(fill = Antibody), stat = 'identity', width = 0.5) +
  geom_text(aes(label = resistant_frac), vjust = -.2) +
  scale_y_continuous(expression(paste("resistance (% IC"[50]>=10,")")),
                     limits = c(0,50), breaks = 0:5*10) +
  scale_x_discrete("", limits = c("10-1074","3BNC117", "VRC07-523-LS")) +
  theme_bw() + theme(
    axis.title.y = element_text(size = 9),
    #text = element_text(size = 11),
    axis.text.x = element_blank(),
    strip.text = element_text(size = 12),
    legend.position = "none",
    plot.margin = unit(c(0.1, .1, -.5, .1), "lines")
  )

ic50_dist = catnap_dat %>%
  filter(clean_IC50 < ic50_cutoff) %>%
  ggplot(aes(x = Antibody, y = clean_IC50, colour = Antibody)) +
  geom_boxplot() +
  scale_x_discrete("", limits = c("10-1074","3BNC117", "VRC07-523-LS")) +
  scale_y_log10(ic50_title,
                           breaks = head(ic50_breaks,-1), labels =  head(ic50_labels, -1)) +
    theme_bw() +
  theme(plot.margin = unit(c(0,0,0.1,0), "lines"), 
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    #text = element_text(size = 11),
    strip.text = element_blank(),
        legend.position = "none") 

ic50_pl = cowplot::plot_grid(res_pl, ic50_dist, nrow = 2, align = "v",
                   rel_heights = c(1.4,2))

ic50_pl

```

## Pegu 

```{r pegu-curve}

pegu_dat = tibble(
  titer = c(91, 219, 685),
  neut = c(0.5, 0.75, 0.95)
)

empirical_pd_fit = drc::drm(neut ~ titer, data = pegu_dat, 
                        fct = drc::LL.5(names = c("slope", "lower", "upper",
                                                  "inflection", "asymmetry"),
                             fixed = c(NA, 0, 1, NA, NA)))

empirical_4pl_fit = drc::drm(neut ~ titer, data = pegu_dat, 
                        fct = drc::LL.4(names = c("slope", "lower", "upper",
                                                  "inflection"),
                             fixed = c(NA, 0, 1, NA)))


pegu_pl = tibble(
  titer = 10^seq(-0.1, 4.5, length = 100),
  invitro = 1 - (1 + titer)^(-1),
  empirical = predict(empirical_pd_fit, data.frame(titer = titer)),
  potency =  1 - (1 + titer/91)^(-1)
) %>%
  gather(model, neut, invitro, empirical, potency) %>%
  ggplot(aes(x = titer, y = neut * 100)) +
  geom_line(aes(colour = model, linetype = model)) +
  geom_point(data = pegu_dat) +
  scale_y_continuous(breaks = 0:10 * 10) +
  labs(y = "empirical protection (%) (Pegu et al.)", 
       colour = "", linetype = "") +
  scale_color_manual(breaks = c("invitro", "empirical", "potency"), 
                     labels = c("In vitro neutralization", "5PL", 
                                "Potency reduction"),
                     values = c("black", "black", "gray")) +
  scale_linetype_manual(breaks = c("invitro", "empirical", "potency"), 
                     labels = c("In vitro neutralization", "5PL", 
                                "Potency reduction"),
                     values = c(2, 1, 1)) +
  scale_x_log10(breaks = 10^(0:4), labels = as.character(10^(0:4))) +
  theme(axis.text.x = element_text(size = 8), 
        legend.key.width = unit(0.5, "cm"), 
        legend.key.height =  unit(0.5, "cm"),
        legend.margin = margin(t = -10,0,0,0),
        legend.box.background = element_rect(colour = "black"),
        legend.position = c(1, 0), legend.text = element_text(size = 7),
        legend.direction = "vertical", 
        legend.justification = c(1.01, -.10))

pegu_pl

```

## AMP

```{r amp-curve}

amp_conc_data = tribble(
  ~dose_mgkg, ~study, ~n, ~peak, ~midpoint, ~trough,
  10, 704, 895, 65.7, 23.1, 5.8,
  30, 704, 894, 212, 55.1, 15.4,
  10, 703, 642, 88.9, 19.8, 4.7,
  30, 703, 645, 257.9, 44.7, 8.3
)

amp_conc_avg =  amp_conc_data %>%
  mutate(weight = n / sum(n)) %>%
  summarize(across(c(peak, midpoint, trough), ~sum(. * weight)), .groups = "drop")

amp_dat_all = read_excel("data/nejm_fig3a.xlsx", col_names = c("ic80", "pe")) %>%
  crossing(time = c("peak", "midpoint", "trough")) %>%
  left_join(gather(amp_conc_avg, time, mean_conc, midpoint, peak, trough), by = "time") %>%
  mutate(
    neut = pe/100,
    id80 = mean_conc/ic80,
    titer = 4 * id80
  ) 

amp_dat = dplyr::filter(amp_dat_all, time == "midpoint")

amp_pd_fit = drc::drm(neut ~ titer, data = amp_dat,
                        fct = drc::LL.5(names = c("slope", "lower", "upper",
                                                  "inflection", "asymmetry"),
                             fixed = c(NA, 0, 1, NA, NA)))

# 370
amp_id50 = drc::ED(amp_pd_fit, 50, display = F)[[1]]
amp_id95 = drc::ED(amp_pd_fit, 95, display = F)[[1]]


amp_pl = tibble(
  titer = 10^seq(-0.1, 4.5, length = 100),
  invitro = 1 - (1 + titer)^(-1),
  empirical = predict(amp_pd_fit, data.frame(titer = titer)),
  potency =  1 - (1 + titer/370)^(-1)
) %>%
  gather(model, neut, invitro, empirical, potency) %>%
  ggplot(aes(x = titer, y = neut * 100)) +
  geom_line(aes(colour = model, linetype = model)) +
  geom_point(data = amp_dat) +
  scale_y_continuous(breaks = 0:10 * 10) +
  labs(y = "empirical protection (%) (AMP)", 
       colour = "", linetype = "") +
  scale_color_manual(breaks = c("invitro", "empirical", "potency"), 
                     labels = c("In vitro neutralization", "5PL", 
                                "Potency reduction"),
                     values = c("black", "black", "gray")) +
  scale_linetype_manual(breaks = c("invitro", "empirical", "potency"), 
                     labels = c("In vitro neutralization", "5PL", 
                                "Potency reduction"),
                     values = c(2, 1, 1)) +
  scale_x_log10(breaks = 10^(0:4), labels = as.character(10^(0:4))) +
  theme(axis.text.x = element_text(size = 8), 
        legend.key.width = unit(0.5, "cm"), 
        legend.key.height =  unit(0.5, "cm"),
        legend.margin = margin(t = -10,0,0,0),
        legend.box.background = element_rect(colour = "black"),
        legend.position = c(1, 0), legend.text = element_text(size = 7),
        legend.direction = "vertical", 
        legend.justification = c(1.01, -.10))

amp_pl

```


```{r amp-curve-id80, eval = F}


amp_dat_id80 = read_excel("data/nejm_fig3a.xlsx", col_names = c("ic80", "pe")) %>%
  mutate(
    neut = pe/100,
    id80 = 19.6/ic80,
    titer = 4 * id80
  ) 


amp_id80_pd_fit = drc::drm(neut ~ id80, data = amp_dat_id80,
                        fct = drc::LL.5(names = c("slope", "lower", "upper",
                                                  "inflection", "asymmetry"),
                             fixed = c(NA, 0, 1, NA, NA)))

# 370
amp_id50 = drc::ED(amp_id80_pd_fit, 50, display = F)[[1]]
amp_id90 = drc::ED(amp_id80_pd_fit, 90, display = F)[[1]]
amp_id95 = drc::ED(amp_id80_pd_fit, 95, display = F)[[1]]


amp_id80_pl = tibble(
  id80 = 10^seq(-0.1, 4.5, length = 100),
  empirical = predict(amp_id80_pd_fit, data.frame(id80 = id80)),
  potency =  1 - (1 + id80/amp_id50)^(-1)
) %>%
  gather(model, neut, empirical, potency) %>%
  ggplot(aes(x = id80, y = neut * 100)) +
  geom_line(aes(colour = model, linetype = model)) +
  geom_point(data = amp_dat) +
  scale_y_continuous(breaks = 0:10 * 10) +
  labs(y = "empirical protection (%) (AMP)", 
       colour = "", linetype = "") +
  scale_color_manual(breaks = c( "empirical", "potency"), 
                     labels = c( "5PL", 
                                "Potency reduction"),
                     values = c("black", "gray")) +
  scale_linetype_manual(breaks = c("empirical", "potency"), 
                     labels = c( "5PL", 
                                "Potency reduction"),
                     values = c(1, 1)) +
  scale_x_log10(breaks = 10^(0:4), labels = as.character(10^(0:4))) +
  theme(axis.text.x = element_text(size = 8), 
        legend.key.width = unit(0.5, "cm"), 
        legend.key.height =  unit(0.5, "cm"),
        legend.margin = margin(t = -10,0,0,0),
        legend.box.background = element_rect(colour = "black"),
        legend.position = c(1, 0), legend.text = element_text(size = 7),
        legend.direction = "vertical", 
        legend.justification = c(1.01, -.10))

amp_id80_pl

```


```{r amp-pegu-dr}

dr_labels = tibble(
  breaks = c("invitro", "pegu_empirical", "pegu_potency", "amp_empirical", "amp_potency"),
  labels =  c("In vitro neutralization", "NHP 5PL", bquote("NHP potency reduction (" *rho*" = 91)"), 
              "AMP 5PL", bquote("AMP potency reduction (" *rho*" = 370)"))
)


amp_pegu_dr = tibble(
  titer = 10^seq(-0.1, 4.5, length = 100),
  invitro = 1 - (1 + titer)^(-1),
  pegu_empirical = predict(empirical_pd_fit, data.frame(titer = titer)),
  pegu_potency =  1 - (1 + titer/91)^(-1),
  amp_empirical = predict(amp_pd_fit, data.frame(titer = titer)),
  amp_potency =  1 - (1 + titer/370)^(-1)
) %>%
  gather(model, neut, invitro:amp_potency) %>%
  ggplot(aes(x = titer, y = neut * 100)) +
  geom_line(aes(colour = model, linetype = model)) +
  geom_point(data = pegu_dat, colour = "blue") +
  geom_point(data = amp_dat, colour = "red") +
  scale_y_continuous(breaks = 0:10 * 10) +
  labs(y = "empirical protection (%)", 
       colour = "", linetype = "") +
  scale_color_manual(breaks = dr_labels$breaks, 
                     labels = dr_labels$labels,
                     values = c("black", "blue", alpha("blue", 1), "red", alpha("red", 1))
                     ) +
  scale_linetype_manual(breaks = dr_labels$breaks, 
                     labels = dr_labels$labels,
                     values = c(2, 1, 3, 1, 3)) +
  scale_x_log10(breaks = 10^(0:4), labels = as.character(10^(0:4))) +
  theme(axis.text.x = element_text(size = 8), 
        legend.key.width = unit(0.5, "cm"), 
        legend.key.height =  unit(0.5, "cm"),
        legend.margin = margin(t = -10,0,0,0),
        legend.box.background = element_rect(colour = "black"),
        legend.position = c(1, 0), legend.text = element_text(size = 7),
        legend.direction = "vertical", 
        legend.justification = c(1.01, -.10))

amp_pegu_dr

```

## Empirical DR

```{r empirical-dr}

conc_range = 10^(seq(-1, 3, length = 50))

empirical_iip = map_df(conc_range, function(i) {
  pd_data %>%
    mutate_if(is.numeric, ~ (-log10(1 - empirical_pd_fun(i / .x, mod5pl = empirical_pd_fit)))) %>%
    gather(analyte, iip, -subtype, -Reference, -Virus) %>%
    group_by(analyte) %>%
    summarize_at(vars(iip), 
                 list(mean = mean,  median = median,
                      min = min, max = max, 
                      lower95 = ~quantile(., .025),
                      upper95 = ~quantile(., 0.975))
                 ) %>%
    mutate(concentration = i,
           outcome = "Pegu est")
}) 

amp_iip = map_df(conc_range, function(i) {
  pd_data %>%
    mutate_if(is.numeric, ~ (-log10(1 - empirical_pd_fun(i / .x, mod5pl = amp_pd_fit)))) %>%
    gather(analyte, iip, -subtype, -Reference, -Virus) %>%
    group_by(analyte) %>%
    summarize_at(vars(iip), 
                 list(mean = mean,  median = median,
                      min = min, max = max, 
                      lower95 = ~quantile(., .025),
                      upper95 = ~quantile(., 0.975))
                 ) %>%
    mutate(concentration = i,
           outcome = "AMP est")
}) 

raw_iip = map_df(conc_range, function(i) {
  pd_data %>%
    mutate_if(is.numeric, ~ (-log10(1 - titer2neut(i / .x)))) %>%
    gather(analyte, iip, -subtype, -Reference, -Virus) %>%
    group_by(analyte) %>%
    summarize_at(vars(iip), 
                 list(mean = mean,  median = median,
                      min = min, max = max, 
                      lower95 = ~quantile(., .025),
                      upper95 = ~quantile(., 0.975))
                 ) %>%
    mutate(concentration = i,
           outcome = "in vitro")
}) 


marginal_iip_pl = empirical_iip %>%
  bind_rows(raw_iip) %>%
  bind_rows(amp_iip) %>%
  ggplot(aes(x = concentration, y = mean, colour = analyte, linetype = outcome)) +
  geom_hline(yintercept = -log10(1 - c(0.5, 0.95, 0.99)), linetype = "dotted", size = 0.25) +
  geom_line() +
  scale_x_log10(breaks = 10^(-1:3), labels = as.character(10^(-1:3))) +
  labs(y = "mean iip", linetype = "") +
  scale_linetype_manual(breaks = c("AMP est", "Pegu est", "in vitro"), values = c(1,4,3)) +
  annotate("text", x = 1000, y = -log10(1 - c(0.5, 0.95, 0.99)), 
           label = c("50%", "95%", "99%"), vjust = 0) +
  theme(legend.box =  "vertical")

marginal_iip_pl

```


```{r empirical-coverage}
#drc::ED(empirical_pd_fit, .5, type = "absolute")
# 91
#drc::ED(empirical_pd_fit, .95, type = "absolute")
# 685

empirical_coverage = map_df(conc_range, function(i) {
  pd_data %>%
    mutate_if(is.numeric, ~ (i / .x)) %>% # titer
    gather(analyte, titer, -subtype, -Reference, -Virus) %>%
    group_by(analyte) %>%
    summarize_at(vars(titer), 
                 list(
                   neut_50 = ~(mean(.x > 1)),
                   pegu_50 = ~(mean(.x > 91)),
                   amp_50 = ~(mean(.x > amp_id50)),
                   neut_95 = ~(mean(.x > neut2titer(.95))),
                   pegu_95 = ~(mean(.x > 685)),
                   amp_95 = ~(mean(.x > amp_id95))
                   )
                 ) %>%
    mutate(concentration = i)
}) %>%
  gather(coverage, prop, neut_50:amp_95) %>%
  separate(coverage, into = c("outcome", "level"), sep = "_") %>%
  mutate(
    prot_lab = factor(outcome,
      levels = c("neut", "pegu", "amp"),
      labels =  c("In vitro neutralization", "In vivo protection (NHP)", "In vivo protection (AMP)")
    ),
    level_lab = case_when(
      level == "50" ~ "% viruses > 50%",
      level == "95" ~ "% viruses > 95%"
    )
  )

marg_cov_pl = empirical_coverage %>%
  ggplot(aes(x = concentration, y = prop * 100, colour = analyte)) +
  geom_step() +
  scale_x_log10(breaks = 10^(-1:3), labels = as.character(10^(-1:3))) +
  labs(y = "", x = expression(paste("concentration (", mu, "g/mL)"))) +
  facet_grid(level_lab~prot_lab, switch = "y") +
  theme(strip.placement = "outside", 
        strip.text.y = element_text(size = 12), strip.background.y = element_blank())

marg_cov_pl

```

# Ratio optimization - PKPD

Use independence model: 1 - prod (1 - x)

```{r pkpd-parms}

models = tibble(
  model = list(empirical_pd_fit, amp_pd_fit),
  label = c("pegu", "amp")
)

dose_ratio_list = seq(0, 1, length = 15)

outcome_labels = tibble(
  levels = c("coverage_50", "coverage_95", "mean_iip"),
  labels = c("% viruses protection > 50%",
             "% viruses protection > 95%",
             "Mean protection IIP")
)

outcome_factor = function(pd_outcome) factor(pd_outcome, levels = outcome_labels$levels,
                                            labels = outcome_labels$labels)

```

## Optimizations

```{r setup-opt-parms}

total_ratio = length(dose_ratio_list)
full_double_analyte_list = tibble(
  mab1 = c(dose_ratio_list, dose_ratio_list, rep(0, total_ratio)),
  mab2 = c(1 - dose_ratio_list, rep(0, total_ratio), 1 - dose_ratio_list),
  mab_set = c(rep("10-1074-T:3BNC117-T", total_ratio),
              rep("3BNC117-T:VRC07-523LS", total_ratio),
              rep("10-1074-T:VRC07-523LS", total_ratio)
              
  )
)

triple_parms = all_pk %>%
  unnest(cols = rate_parms) %>%
  dplyr::filter(analyte %in% c("10-1074-T", "3BNC117-T", "VRC07-523LS"))

```

### 2-bnab combination optimization

```{r full-combo-search, eval = opt_2bnab_flag}

# this is the original Pegu runs
combo_pkpd_full = map_df(1:nrow(full_double_analyte_list), function(i_row) {
  
  i = full_double_analyte_list$mab1[i_row]
  j = full_double_analyte_list$mab2[i_row]
  analyte_set = full_double_analyte_list$mab_set[i_row]
  
  pk_setup = triple_parms %>%
    mutate(dose = if_else(
      analyte == "3BNC117-T",
      600 * i,
      if_else(analyte == "10-1074-T", 600 * j,
              600 * (1 - (i + j)))
    )) %>%
    check_dose_calc(total_dose = 600)
  
  active_mabs = filter(pk_setup, dose > 0)[['analyte']]
  total_active = length(active_mabs)
  pk_setup %>%
    create_pk(pk_window) %>%
    pivot_wider(values_from = conc, names_from = analyte) %>%
    rename(mab_3bnc = `3BNC117-T`,
           mab_1074 = `10-1074-T`,
           mab_vrc07 = `VRC07-523LS`) %>%
    group_by(time) %>%
    nest() %>%
    triple_pd_calc(pd_data, mod5pl = empirical_pd_fit) %>%
    select(-data,-protection_res) %>%
    unnest(neut_summary) %>%
    ungroup() %>%
    mutate(
      total_dose = 600,
      total_active = total_active,
      analyte_set = analyte_set,
      active_mabs = paste(active_mabs, collapse = ":"),
      ratio_3bnc = i,
      ratio_1074 = j,
      ratio_vrc = 1 - (i + j)
    )
})  %>%
  mutate(total_active_check = rowSums(cbind(ratio_3bnc > 0, ratio_1074 > 0, ratio_vrc > 0)))
with(combo_pkpd_full, ftable(active_mabs, analyte_set))

# this is the AMP runs, this is coded redundantly because it meets a paper revision request
combo_pkpd_amp = map_df(1:nrow(full_double_analyte_list), function(i_row) {
  
  i = full_double_analyte_list$mab1[i_row]
  j = full_double_analyte_list$mab2[i_row]
  analyte_set = full_double_analyte_list$mab_set[i_row]
  
  pk_setup = triple_parms %>%
    mutate(dose = if_else(
      analyte == "3BNC117-T",
      600 * i,
      if_else(analyte == "10-1074-T", 600 * j,
              600 * (1 - (i + j)))
    )) %>%
    check_dose_calc(total_dose = 600)
  
  active_mabs = filter(pk_setup, dose > 0)[['analyte']]
  total_active = length(active_mabs)
  pk_setup %>%
    create_pk(pk_window) %>%
    pivot_wider(values_from = conc, names_from = analyte) %>%
    rename(mab_3bnc = `3BNC117-T`,
           mab_1074 = `10-1074-T`,
           mab_vrc07 = `VRC07-523LS`) %>%
    group_by(time) %>%
    nest() %>%
    triple_pd_calc(pd_data, mod5pl = amp_pd_fit) %>%
    select(-data,-protection_res) %>%
    unnest(neut_summary) %>%
    ungroup() %>%
    mutate(
      total_dose = 600,
      total_active = total_active,
      analyte_set = analyte_set,
      active_mabs = paste(active_mabs, collapse = ":"),
      ratio_3bnc = i,
      ratio_1074 = j,
      ratio_vrc = 1 - (i + j)
    )
})  %>%
  mutate(total_active_check = rowSums(cbind(ratio_3bnc > 0, ratio_1074 > 0, ratio_vrc > 0)))


save(combo_pkpd_full, file = "output/double-opt.rda")
save(combo_pkpd_amp, file = "output/double-opt-amp.rda")

```

```{r load-dbl-opt, eval = !opt_2bnab_flag}
load("output/double-opt.rda")
load("output/double-opt-amp.rda")

combo_pkpd_all = mutate(combo_pkpd_full, model = "pegu") %>%
  bind_rows(mutate(combo_pkpd_amp, model = "amp"))

```


```{r triple-fixed-ratios}

# ratio1 = 3bnc share
# ratio2 = 10-1074 share 
# vrc07 is remaining

# we will explicitly look at the boundaries
#triple_ratio_grid = head(dose_ratio_list, -1)[-1]
#total_ratios = length(triple_ratio_grid)

# one mab
single_analyte_list = tribble(
  ~mab1, ~mab2,
  1, 0,
  0, 1,
  0, 0
)

# two mab
# mab1 = dose_ratio_list, mab2 = 1 - dose_ratio_list -> mab3 = 0
# mab1 = dose_ratio_list, mab2 = 0 -> mab3 = 1- dose list
# mab1 = 0, mab2 = dose list-> mab3 = 1- dose list

double_analyte_list = tibble(
  mab1 = c(0.5, 0.5, 0),
  mab2 = c(0.5, 0, 0.5)
)

# triple mab
#triple_sample_raw = NULL #lhs::improvedLHS(25, 3)
#triple_sample_norm = NULL #triple_sample_raw/rowSums(triple_sample_raw)

triple_analyte_list = tibble(
  mab1 = c(0.5, 0.25, 0.25, 1/3),
  mab2 = c(0.25, 0.5, 0.25, 1/3)
)
#all(rowSums(triple_analyte_list) < 1)

triple_ratio_list = single_analyte_list %>%
  bind_rows(double_analyte_list) %>%
  bind_rows(triple_analyte_list)

```


```{r triple_fixed, eval = T}

mab_assign = tibble(
  remove_mab = 1:3,
  mabs = c("3BNC117-T", "10-1074-T", "VRC07-523LS"),
  mabs_pl =  c("3BNC117-T", "10-1074-T", "VRC07\n-523LS")
)

triple_pkpd_fixed = map2_df(models$model, models$label, function(mod, mod_lab){
  
  map2_df(triple_ratio_list$mab1, triple_ratio_list$mab2, function(i, j) {
  
  pk_setup = triple_parms %>%
    mutate(dose = if_else(
      analyte == "3BNC117-T",
      600 * i,
      if_else(analyte == "10-1074-T", 600 * j,
              600 * (1 - (i + j)))
    )) %>%
    check_dose_calc(total_dose = 600)
  
  active_mabs = filter(pk_setup, dose > 0)[['analyte']]
  total_active = length(active_mabs)
  pk_setup %>%
    create_pk(pk_window) %>%
    pivot_wider(values_from = conc, names_from = analyte) %>%
    rename(mab_3bnc = `3BNC117-T`,
           mab_1074 = `10-1074-T`,
           mab_vrc07 = `VRC07-523LS`) %>%
    group_by(time) %>%
    nest() %>%
    triple_pd_calc(pd_data, mod5pl = mod) %>%
    select(-data,-protection_res) %>%
    unnest(neut_summary) %>%
    ungroup() %>%
    mutate(
      total_dose = 600,
      total_active = total_active,
      active_mabs = paste(active_mabs, collapse = ":"),
      ratio_3bnc = i,
      ratio_1074 = j,
      ratio_vrc = 1 - (i + j)
    )
}) %>%
  mutate(
    total_active_check = rowSums(cbind(ratio_3bnc > 0, ratio_1074 > 0, ratio_vrc > 0)),
    model = mod_lab
    )
})


```


```{r combo-opts, eval = opt_3bnab_flag}

opt_outcomes = crossing(
  outcomes = c("coverage_50", "coverage_95"),
  endpoints = c("trough", "auc")
)

# Pegu

combo_opt = map2_df(opt_outcomes$outcomes, opt_outcomes$endpoints, function(out, end){
  map_df(1:3, function(i){
    opt = optimize(combo_opt_fun,
      interval = c(-10, 10),
      pk_parms = triple_parms,
      total_dose = 600,
      pk_window = pk_window,
      pd_data = pd_data,
      outcome_var = out,
      pk_endpoint = end,
      mod5pl = empirical_pd_fit,
      remove_mab = i
    )
    #browser()
    opt_ratios = organize_opt_ratios(calc_opt_ratios(opt$minimum), remove_mab = i)
    
    tibble(
      outcome = out,
      trough_time = max(pk_window),
      endpoint = end,
      analytes = paste0(mab_assign$mabs[-i], collapse = ":"),
      ratio_3bnc = opt_ratios[1],
      ratio_1074 = opt_ratios[2],
      ratio_vrc = opt_ratios[3],
      value = -opt$objective
    )
    
  })
})

triple_opt = map2_df(opt_outcomes$outcomes, opt_outcomes$endpoints, function(out, end){
  opt = optim(c(-2.381427, 0.705), combo_opt_fun, 
              pk_parms = triple_parms, total_dose = 600, pk_window = pk_window,
              pd_data = pd_data, outcome_var = out, pk_endpoint = end, mod5pl = empirical_pd_fit
)
  #browser()
  opt_ratios = calc_opt_ratios(opt$par)
  
  tibble(
    outcome = out,
    trough_time = max(pk_window),
    endpoint = end,
    ratio_3bnc = opt_ratios[1],
    ratio_1074 = opt_ratios[2],
    ratio_vrc = opt_ratios[3],
    value = -opt$value
  )
  
})

# AMP, structure follows revision edits

combo_opt_amp = map2_df(opt_outcomes$outcomes, opt_outcomes$endpoints, function(out, end){
  map_df(1:3, function(i){
    opt = optimize(combo_opt_fun,
      interval = c(-10, 10),
      pk_parms = triple_parms,
      total_dose = 600,
      pk_window = pk_window,
      pd_data = pd_data,
      outcome_var = out,
      pk_endpoint = end,
      mod5pl = amp_pd_fit,
      remove_mab = i
    )
    #browser()
    opt_ratios = organize_opt_ratios(calc_opt_ratios(opt$minimum), remove_mab = i)
    
    tibble(
      outcome = out,
      trough_time = max(pk_window),
      endpoint = end,
      analytes = paste0(mab_assign$mabs[-i], collapse = ":"),
      ratio_3bnc = opt_ratios[1],
      ratio_1074 = opt_ratios[2],
      ratio_vrc = opt_ratios[3],
      value = -opt$objective
    )
    
  })
})

triple_opt_amp = map2_df(opt_outcomes$outcomes, opt_outcomes$endpoints, function(out, end){
  opt = optim(c(-2.381427, 0.705), combo_opt_fun, 
              pk_parms = triple_parms, total_dose = 600, pk_window = pk_window,
              pd_data = pd_data, outcome_var = out, pk_endpoint = end,  mod5pl = amp_pd_fit)
  #browser()
  opt_ratios = calc_opt_ratios(opt$par)
  
  tibble(
    outcome = out,
    trough_time = max(pk_window),
    endpoint = end,
    ratio_3bnc = opt_ratios[1],
    ratio_1074 = opt_ratios[2],
    ratio_vrc = opt_ratios[3],
    value = -opt$value
  )
  
})


save(combo_opt, triple_opt, file = "output/triple-opt.rda")
save(combo_opt_amp, triple_opt_amp, file = "output/triple-opt-amp.rda")

```

```{r load-triple-opt, eval = !opt_3bnab_flag}

load("output/triple-opt.rda")
load("output/triple-opt-amp.rda")

all_triple_opt = triple_opt %>%
  mutate(model = "pegu", combo = "3-mab") %>%
  bind_rows(mutate(triple_opt_amp, model = "amp", combo = "3-mab"))


all_opt = all_triple_opt %>%
  bind_rows(mutate(combo_opt, model = "pegu", combo = "2-mab")) %>%
  bind_rows(mutate(combo_opt_amp, model = "amp", combo = "2-mab"))
  


```

### Triple optimal ratios

```{r opt-res-table}

all_opt %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  dplyr::select(combo, endpoint, outcome, model, everything(), -trough_time, -analytes) %>%
  mutate(outcome = paste0(str_remove(outcome, "coverage_"), "%"),
         value = round(value * 100)
         ) %>%
  mutate_at(vars(contains("ratio")), round, digits = 2) %>%
  rename(`% viruses` = value) %>%
  arrange(combo, outcome, endpoint, model) %>%
  dplyr::select(-combo) %>%
  write_csv("output/empirical-opt-ratios.csv") %>%
  kable(caption = "optimal ratios") %>%
  kable_styling(full_width = F) %>%
  pack_rows("Two mAb", 1, 24) %>%
  pack_rows("Three mAb", 25, 32)

```

### Single mab results

```{r triple-marg-process}

marg_res = triple_pkpd_fixed %>%
  filter(total_active == 1) %>%
  gather(outcome, value, mean_iip, coverage_50, coverage_95) %>%
  gather(analyte, ratio, contains("ratio")) %>%
  dplyr::filter(ratio == 1) %>%
  separate(analyte, into = c("tmp", "ratio_cat")) %>%
  group_by(outcome, ratio_cat, model) %>%
  arrange(time) %>%
  summarize(
    trough_time = max(time),
    trough = value[which.max(time)],
    auc = pracma::trapz(time, value)/trough_time
    ,.groups = "drop"
  ) %>%
  gather(endpoint, value, trough, auc) %>%
  mutate(
    ratio_cat = case_when(
      ratio_cat == "3bnc" ~ mab_assign$mabs_pl[1], 
      ratio_cat == "1074" ~ mab_assign$mabs_pl[2], 
      ratio_cat == "vrc" ~ mab_assign$mabs_pl[3]
      ),
    value = if_else(str_detect(outcome, "coverage"), value * 100, value),
    outcome2 = outcome_factor(outcome),
    category = "Single mAb"
    )

marg_res %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  ggplot(aes(x = ratio_cat, y = value, colour = outcome2, shape = endpoint)) +
  geom_point() +
  facet_wrap(~model)

```

### Combination results

```{r full-combo-ratios}

#unique(combo_pkpd_full$total_active)
# unique(combo_pkpd_full$active_mabs)
# with(combo_pkpd_full, ftable(active_mabs))

full_combo_res = combo_pkpd_all %>%
  gather(outcome, value, mean_iip, coverage_50, coverage_95) %>%
  group_by_at(vars(outcome, analyte_set, model, contains("ratio"))) %>%
  arrange(time) %>%
  summarize(
    max_ratio = max(c(ratio_3bnc, ratio_1074, ratio_vrc)),
    trough_time = max(time),
    trough = value[which.max(time)],
    auc = pracma::trapz(time, value)/trough_time
    , .groups = "drop"
  ) %>%
  gather(endpoint, value, trough, auc) %>%
  mutate(
    ratio = if_else(str_split_fixed(analyte_set, ":", n = 2)[, 2] == "3BNC117-T",
                    ratio_3bnc, ratio_vrc),
    value = if_else(str_detect(outcome, "coverage"), value * 100, value)
  ) %>%
  rename(analytes = analyte_set)
  

full_combo_res %>%
  dplyr::filter(str_detect(outcome, "coverage"))   %>%
  ggplot(aes(x = ratio, y = value, colour = outcome, shape = endpoint, linetype = model)) +
  geom_line() +
  facet_grid(endpoint ~ analytes)

```

```{r triple-2comb-process}

combo_res_summary = triple_pkpd_fixed %>%
  filter(total_active == 2) %>%
  gather(outcome, value, mean_iip, coverage_50, coverage_95) %>%
  group_by_at(vars(model, outcome, active_mabs, contains("ratio"))) %>%
  arrange(time) %>%
  summarize(
    max_ratio = max(c(ratio_3bnc, ratio_1074, ratio_vrc)),
    trough_time = max(time),
    trough = value[which.max(time)],
    auc = pracma::trapz(time, value)/trough_time,
    .groups = "drop"
  ) %>%
  gather(endpoint, value, trough, auc) 

combo_res_opt = combo_opt %>%
  rename(category = analytes) %>%
  mutate(
    ratio_cat = "optimal",
    #ratio_cat = paste0("best\n", str_replace(analytes, ":", "\n")),
    value = if_else(str_detect(outcome, "coverage"), value * 100, value),
    outcome2 = outcome_factor(outcome),
    analytes = if_else(category == "3BNC117-T:10-1074-T", "10-1074-T:3BNC117-T",
                            category)
    )

combo_res_best = full_combo_res %>%
  group_by(analytes, model, outcome, endpoint) %>%
  summarize(
    ratio_cat = "optimal",
    ratio = ratio[which.max(value)],
    value = max(value),
    .groups = "drop"
    ) %>%
  mutate(
    category = if_else(
      analytes == "10-1074-T:3BNC117-T",
      "3BNC117-T:10-1074-T",
      analytes
    )
  )

combo_res_fixed = combo_res_summary %>%
  dplyr::filter(max_ratio == 0.5) %>%
  mutate(
    ratio_cat = "1:1",
    #ratio_cat = paste0("1:1\n", str_replace(active_mabs, ":", "\n")),
    value = if_else(str_detect(outcome, "coverage"), value * 100, value),
    category = if_else(
      active_mabs == "10-1074-T:3BNC117-T",
      "3BNC117-T:10-1074-T",
      active_mabs
    )
  ) %>%
  rename(analytes = active_mabs)

all_combo = combo_res_fixed %>%
  mutate(
    ratio = if_else(
      str_split_fixed(analytes, ":", n = 2)[, 2] == "3BNC117-T",
      ratio_3bnc,
      ratio_vrc
    )
  ) %>%
  bind_rows(combo_res_best) %>%
  mutate(category = str_replace(category, ":", ":\n"),
         outcome2 = outcome_factor(outcome)
)

all_combo %>%
  dplyr::filter(str_detect(outcome, "coverage"))   %>%
  ggplot(aes(x = ratio_cat, y = value, colour = outcome2, shape = endpoint)) +
  geom_point() +
  facet_grid(model~category)

```


```{r combo-full-opt}

full_combo_res %>%
  dplyr::filter(str_detect(outcome, "coverage"))   %>%
  ggplot(aes(x = ratio, y = value, colour = endpoint, linetype = as.factor(parse_number(outcome)))) +
  geom_line() + 
  geom_point(data = dplyr::filter(all_combo, str_detect(outcome, "coverage")), aes(x = ratio, shape = ratio_cat)) +
  facet_grid(model~analytes) +
  labs(colour  ="coverage_pct") +
  theme(legend.position = "right", legend.title = element_text(size= 8), legend.text = element_text(size= 8))

```

```{r triple-res-process}

triple_res_summary = triple_pkpd_fixed %>%
  filter(total_active == 3) %>%
  gather(outcome, value, mean_iip, coverage_50, coverage_95) %>%
  group_by_at(vars(model, outcome, active_mabs, contains("ratio"))) %>%
  arrange(time) %>%
  summarize(
    min_ratio = min(c(ratio_3bnc, ratio_1074, ratio_vrc)),
    max_ratio = max(c(ratio_3bnc, ratio_1074, ratio_vrc)),
    trough_time = max(time),
    trough = value[which.max(time)],
    auc = pracma::trapz(time, value)/trough_time,
    .groups = "drop"
  ) %>%
  gather(endpoint, value, trough, auc) 

triple_res_fixed = triple_res_summary %>%
  mutate(
    ratio_cat = if_else(min_ratio == 1/3, "1:1:1", 
                        paste(ratio_3bnc*4, ratio_1074*4, ratio_vrc*4, sep = ":")),
    value = if_else(str_detect(outcome, "coverage"), value * 100, value),
    outcome2 = outcome_factor(outcome)
    ) %>%
  mutate(category = "3BNC117-T:10-1074-T:VRC07-523LS")

triple_res_opt = all_triple_opt %>%
  mutate(
    ratio_cat = "optimal",
    value = if_else(str_detect(outcome, "coverage"), value * 100, value),
    outcome2 = outcome_factor(outcome)
  )%>%
  mutate(category = "3BNC117-T:10-1074-T:VRC07-523LS")

triple_res_fixed %>%
  bind_rows(triple_res_opt) %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  ggplot(aes(x = ratio_cat, y = value, colour = endpoint, shape = outcome2)) +
  geom_point() +
  facet_grid(model~category)
  

```

### Manuscript figures

```{r orig-full-opt-pl, fig.width = 8}

succinct_main_fig = marg_res %>%
  bind_rows(all_combo) %>%
  bind_rows(triple_res_fixed) %>%
  bind_rows(triple_res_opt) %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  mutate(category = fct_relevel(category, "Single mAb"),
         category = fct_relevel(category, "3BNC117-T:10-1074-T:VRC07-523LS", after = Inf)) %>%
  ggplot(aes(x = ratio_cat, y = value, colour = outcome2, shape = endpoint)) +
  geom_point(size = 2.5) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", colour = "", shape = "", y = "% of viruses") +
  facet_grid(cols = vars(category), rows = vars(model), scale = "free_x", space = "free_x", switch = "x") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        legend.direction = "vertical", axis.text.x = element_text(size = 8))

succinct_main_fig


```


```{r dose-response-fig,  fig.height=6, fig.width=6}

amp_pegu_dr

```

```{r empirical-paper-marginal, fig.height=9, fig.width=7.5}

top_row = plot_grid(man_pk_pl, ic50_pl + theme(axis.text.x = element_text()), 
                    align = "h", axis = "t",  nrow = 1,
                    rel_widths = c(1, 1),
                    labels = c("A", "B"), vjust = 1, hjust = 0)

plot_grid(top_row, marg_cov_pl + theme(legend.position = "none"), labels = c("", "C"), 
          rel_heights = c(1.75, 2), nrow = 2)

```


```{r empirical-paper-assembly, fig.width=7.5}

yaxis_off = theme(axis.title.y = element_blank(),
                  axis.text.y = element_blank(),
                  strip.text.y = element_blank()
)

combo_best_pl = dplyr::filter(combo_res_best, str_detect(outcome, "coverage")) %>%
   mutate(analytes2 = str_replace_all(analytes, ":", ":\n"))

single_mab_pl = marg_res %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  mutate(endpoint2 = paste(endpoint,"(% viruses)")) %>%
  ggplot(aes(x = ratio_cat, y = value, colour = model)) +
  geom_point(size = 2, aes(group = outcome2, shape = outcome2), position = position_dodge(width = 0.4)) +
  scale_y_continuous(limits = c(0, 100), breaks = 0:4*25) +
  labs(x = "", shape = "", y = "% of viruses") +
  scale_color_discrete("protection correlate", breaks = c("pegu", "amp"),
                       labels = c("NHP (Pegu et al.)", "AMP (Corey et al.)")) +
  scale_shape_manual("PKPD endpoint", values = c(16, 1)) +
  facet_grid(cols = vars(category), rows = vars(endpoint2), scale = "free_x", space = "free_x", 
             switch = "both") +
    theme(strip.placement = "outside", strip.background = element_blank(),
        legend.direction = "horizontal", legend.box = "vertical",
        legend.spacing.y = unit(-0.75, "lines"), legend.box.margin = margin(b = 5, t = 2),
        legend.background = element_blank(),
        axis.text.x = element_text(size = 6), strip.text.y = element_text(size = 8)) +
  labs(x = "", y = "") 

pl_leg = get_legend(single_mab_pl)
single_mab_pl 

combo_mab_pl = full_combo_res %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  mutate(analytes2 = str_replace_all(analytes, ":", ":\n")) %>%
  ggplot(aes(x = ratio, y = value, colour = model)) +
  geom_line(aes(linetype = outcome)) + 
  scale_y_continuous(limits = c(0, 100), breaks = 0:4*25) +
  geom_point(data = combo_best_pl, aes(x = ratio, shape = outcome), size = 2) +
  scale_shape_manual(values = c(16, 1)) +
  facet_grid(cols = vars(analytes2), rows = vars(endpoint), 
             scale = "free_x", space = "free_x", switch = "x") +
  scale_x_continuous(breaks = 0.5, labels = "1:1") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        legend.position = "none", axis.text.x = element_text(size = 8)) +
  labs(x = "", y = "") 

combo_mab_pl + theme(legend.position = "top", legend.box = "vertical")

triple_mab_pl = triple_res_opt %>%
  bind_rows(triple_res_fixed) %>%
  mutate(category2 = stringi::stri_replace(category, fixed = ":", ":\n", mode = "last")) %>%
  dplyr::filter(str_detect(outcome, "coverage")) %>%
  ggplot(aes(x = ratio_cat, y = value, colour = model)) +
  geom_point(size = 2.5, aes(shape = outcome, group = outcome), position= position_dodge(width = 0.4)) +
  scale_shape_manual("PKPD endpoint", values = c(16, 1)) +
  scale_y_continuous(limits = c(0, 100), breaks = 0:4*25) +
  labs(x = "", colour = "", shape = "", y = "% of viruses") +
  facet_grid(cols = vars(category2), rows = vars(endpoint), 
             scale = "free_x", space = "free_x", switch = "x") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        legend.position = "none", axis.text.x = element_text(size = 8)) +
  labs(x = "") 

triple_mab_pl + theme(legend.position = "top", legend.box = "vertical")

```

```{r empirical-paper-pl, fig.width=7.5}

combo_pl = plot_grid(single_mab_pl + theme(legend.position = "none"), 
          combo_mab_pl + yaxis_off, triple_mab_pl + yaxis_off,
          axis = "tb",
          align = "h", rel_widths = c(0.9, 1.1, 1), nrow = 1)

plot_grid(pl_leg, combo_pl, nrow = 2, rel_heights = c(1,12))

```

```{r fig5-eps, fig.width=7.5, dev = "postscript", include = F}

plot_grid(pl_leg, combo_pl, nrow = 2, rel_heights = c(1,12))

```
