---
title: "PKPD analysis functions"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    code_folding: none
editor_options:
  chunk_output_type: console
---


## Pharmacodynamic functions (`neutralization-funs.R`)

This suite of functions performs calculations to convert between titer, neutralization, and IIP; and to perform Bliss-Hill independence calculations.

```{r, code=readLines("../code/neutralization-funs.R"), eval = F}
```


## 1-cmpt ratio optimization functions (`onecmptPKPD-sims.R`)

This suite of functions performs PKPD simulations for varying PK/PD inputs sampling viruses accordingly assuming a 1-cmpt PK model. These functions were written to have generalized utility. Functions from here were also used for the [simulation study](https://github.com/bryanmayer/pkpd-bnab-project/blob/main/code/simulation-batch.R) and the bi-specific synergy simulations.

```{r, code=readLines("../code/onecmptPKPD-sims.R"), eval = F}
```

### Processing 1-cmpt theoretical results - [See here](process-sims.html)

## 3-bnab case study ratio optimization functions (`empirical-opt-funs.R`)

These functions follow similar structure to 1-cmpt but were specifically curated for this 3-bnab case study. They expect labels for 3BNC117, 10-1074, and VRC07-523LS; perform 2-cmpt PK simulations; and utilize titer-protection dose-response 5PL model relationships fit as drc objects.

```{r, code=readLines("../code/empirical-opt-funs.R"), eval = F}
```

### workflowr functions

## Build project (`build_all.R`)

Publishes the project website by compiling all of the Rmarkdown files using the workflowr API.

```{r, code=readLines("../code/build_all.R"), eval = F}
```

## Publish project (`publish_all.R`)

Publishes the project website by compiling all of the Rmarkdown files using the workflowr API.

```{r, code=readLines("../code/publish_all.R"), eval = F}
```


