---
title: "Sensitivity correlation analysis code (python)"
author: "Bryan Mayer, Daniel Reeves (python)"
date: "3/14/2022"
output:
  workflowr::wflow_html:
    code_folding: none
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

The sensitivity analysis of the theoretical simulations was performed using python jupyter notebooks. To display the code here, the notebooks were converted into Rmd. 

```{r, eval = F}
rmarkdown:::convert_ipynb("code/python/correlations.ipynb")
rmarkdown:::convert_ipynb("code/python/correlations-optimal.ipynb")
```

The code echoed here is still clunky and the notebooks are best viewed using `jupyter notebook` within the `code/python/` directory.

https://github.com/bryanmayer/pkpd-bnab-project/tree/main/code/python


## PKPD outcome sensitivity (Figure 2) (`correlations.ipynb`)

```{python code=readLines("../code/python/correlations.Rmd"), eval = F}
```

## Optimal ratio sensitivity (Figure 3) (`correlations-optimal.ipynb`)

```{python, code=readLines("../code/python/correlations-optimal.Rmd"), eval = F}
```

