---
title: " Combination titer analysis"
author: "Bryan Mayer"
date: "2021-07-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, message = F, warning = F}
library(metR)
library(tidyverse)
library(kableExtra)
library(cowplot)
library(here)
library(scales)
library(viridis)
source(here("code/neutralization-funs.R"))
source(here("code/onecmptPKPD-sims.R"))
theme_set(theme_bw() + theme(legend.position = "top"))
```

```{r setup}

paste_ic50s = function(x) paste0("{",paste(x, collapse = ", "), "}")
conc_range = 10^(seq(-2, 2.5,  length = 50))
conc_grid = crossing(a = conc_range, b = conc_range)

```

```{r funs}

create_virus_pkpd = function(virus_ic50s, virus_name = NA_character_,
                             pk_data_in){
  pk_data_in %>%
    pivot_wider(names_from = "mab", values_from = "conc") %>%
    rowwise() %>%
    mutate(
      virus = virus_name,
      ID501 = bNAb1 / virus_ic50s[1],
      ID502 = bNAb2 / virus_ic50s[2],
      additivity_ID50 = ID501 + ID502,
      additivity_neut = 1 - 1/(1+ additivity_ID50),
      BH_ID50 = calc_BHID50(ID501, ID502),
      BH_neut = titer2BH(ID501, ID502)
    ) 
}

```

## Supplementary example: total concentration vs. titer vs. IIP by interaction model


From earlier iteration, we want concentration ratios of 1:1 and 1:10; and
\{1, 0.1\} IC50s.

```{r simple-setup}

titer_title = expression(paste("Predicted experimental ID"[50]))
ratio_title = "bNAb concentration ratio (bNAB1:bNAb2)"
conc_title = "[Total concentration]"
iip_title = "Neutralization IIP"
ic50_title = "Viral IC50\n{bNAb1, bNAb2}"

conc_labeller = tibble(
  breaks = c(0.1, 1, 10, 100),
  labels = as.character(c(0.1, 1, 10, 100))
)

#supp_IC50_list = list(c(0.05, 0.5), c(5, 0.5))
#supp_IC50_list = list(c(0.01, 0.5), c(1, 0.1))
supp_IC50_list = list(c(1, 0.1))
conc_range = 10^(seq(-1, 2.5,  length = 200))

```


```{r supp-sim}

single_virus_example =  map_df(supp_IC50_list, function(y) {
  purrr::map_df(c(1, 10), function(x) {
    tmp_pk = tibble(
      ic50s = paste_ic50s(y),
      index = 1:length(conc_range),
      bNAb1 = conc_range / (1 + x),
      bNAb2 = conc_range / (1 + 1 / x)
    ) %>%
      gather(mab, conc, bNAb1:bNAb2)
    create_virus_pkpd(virus_ic50s = y, pk_data_in = tmp_pk) %>%
      mutate(ratio = x)
  })
}) %>% 
  mutate(ratio_lab = factor(ratio, levels = c(1, 10),
                         labels = c( "1:1", "1:10")))

single_virus_example_long = single_virus_example %>%
  dplyr::select(-virus) %>%
    gather(endpoint, value, -index, -ratio, -ic50s,
           -ratio_lab, -ID501, -ID502, -bNAb1, -bNAb2) %>%
    separate(endpoint, into = c("interaction", "endpoint")) %>%
    spread(endpoint, value)

```

### Total concentration vs. titer

Total concentration and experimental titer maintain a consistent relationship across concentration ratios for a given IC50 set.

```{r simple-example-titer}

id50_pred_pl = single_virus_example_long %>%
  mutate(
    interaction = interaction_labeller(interaction)
  ) %>%
  ggplot(aes(
    x = bNAb1 + bNAb2,
    y = ID50
  )) +
  geom_line(aes(color = ratio_lab)) +
  scale_x_log10(breaks = conc_labeller$breaks, 
                labels = conc_labeller$labels) +
  scale_y_log10() +
  coord_cartesian(ylim = c(1, 1000), xlim = c(0.1, 100)) +
  facet_grid(cols = vars(interaction),
             switch = "y") +
  labs(x = conc_title, y = titer_title,
       colour = ratio_title) +
  theme(legend.position = "top", strip.placement = "outside") 

# leg = get_legend(id50_pred_pl + 
#                    guides(colour = guide_legend(title.position = "left"),
#                           linetype = guide_legend(title.position = "left")) +
#                    theme(legend.direction = "vertical", legend.box.margin = margin(b=-2),
#                          legend.background = element_rect(colour = "black", size = 0.1))) 

leg = get_legend(id50_pred_pl)

plot_grid(leg, id50_pred_pl+theme(legend.position = "none"), rel_heights = c(2,12), nrow = 2)

```

### Total concentration vs. neutralization

Total concentration and experimental neutralization have a complicated relationship with BH independence. The benefit of multiplicative independence is realized at high concentrations and balanced ratios, and is not captured by titer in the above plot.


```{r simple-example-neut}

neut_pred_pl = single_virus_example_long %>%
  mutate(
    interaction = interaction_labeller(interaction)
  ) %>%
  ggplot(aes(
    x = bNAb1 + bNAb2,
    y = -log10(1-neut)
  )) +
  geom_line(aes(color = ratio_lab, linetype = ic50s)) +
  scale_x_log10(breaks = conc_labeller$breaks, labels = conc_labeller$labels) +
  coord_cartesian(xlim = c(0.1, 100)) +
  facet_grid(cols = vars(interaction),
             switch = "y") +
  labs(x = conc_title, y = iip_title,
       colour = ratio_title, linetype = ic50_title) +
  theme(legend.position = "none", strip.placement = "outside") 

neut_pred_pl

```

### titer vs. neutralization

Linking the results from above, the relationship between ID50 and IIP under Bliss-Hill independence is not uniquely defined. Different conclusions about optimal ratio can be reached depending on the metric. 

```{r simple-titer-iip}

titer_iip_pl_dat = single_virus_example_long %>%
  dplyr::select(interaction, ID50, neut) %>%
  distinct() %>%
  mutate(
    interaction = interaction_labeller(interaction),
    neut = titer2neut(ID50)
  ) 
  
iip_titer_pl = single_virus_example_long %>%
  mutate(
    interaction = interaction_labeller(interaction)
  ) %>%
  ggplot(aes(
    x = ID50,
    y = -log10(1-neut),
  )) +
  geom_line(aes(color = ratio_lab, linetype = ic50s)) +
  geom_line(data = titer_iip_pl_dat, linetype = "solid") +
  scale_x_log10() +
  facet_grid(cols = vars(interaction),
             switch = "y") +
  coord_cartesian(xlim = c(1, 1000)) +
    labs( x = titer_title, y = iip_title,
       colour = ratio_title, linetype = ic50_title) +
  theme(legend.position = "none", strip.placement = "outside") 

iip_titer_pl

```

### Supplementary figure - combining the results

```{r supp-pl-titer, fig.height=8, fig.width=6}

main_figs = plot_grid(id50_pred_pl+theme(legend.position ="none"),
                                         neut_pred_pl, iip_titer_pl, nrow = 3,
                      align = "v", labels = LETTERS[1:3], label_y = 1.05)

plot_grid(leg, main_figs, nrow = 2, rel_heights = c(1.5,12))                      
                      
```


## Example 2: Two mAb vs. single virus

Comparing additivity and BH highlights non-uniqueness of neutralization and experimental titer with BH. Here, we take use slightly different IC50s and plot combined ID50 vs. neutralization.

```{r single-virus-ic50s}

conc_range = 10^(seq(-2, 2.5,  length = 50))
IC50_list = list(c(1e-2, 1e-2),
                 c(1e-2, 1e-1),
                 c(0.1, 0.1),
                 c(1, .1))

single_virus_example2 = purrr::map_df(c(0.1, 1, 10), function(x) {
  map_df(IC50_list, function(y){
  tibble(input_conc = conc_range,
         ratioAB = x,
         IC50lab = paste_ic50s(y)
         ) %>%
    rowwise() %>%
    mutate(
      ID501 = input_conc/(1 + x) / y[1],
      ID502 = input_conc/(1 + 1/x) / y[2],
      additivity_ID50 = ID501 + ID502,
      additivity_neut = 1 - 1/(1+ additivity_ID50),
      BH_ID50 = calc_BHID50(ID501, ID502),
      BH_neut = titer2BH(ID501, ID502)
    ) %>%
    gather(endpoint, value, -input_conc, -ratioAB, -ID501,-ID502, -IC50lab) %>%
    separate(endpoint, into = c("interaction", "endpoint")) %>%
    spread(endpoint, value)
})
})
titer_iip_pl = tibble(
   ID50 = unique(single_virus_example2$ID50),
   neut = titer2neut(ID50)
 )
 
```


```{r single-virus-pl, fig.height=8, warning=F}

single_virus_example2 %>%
    mutate(
      interaction = paste(interaction_labeller(interaction), "ID50"),
      ratio_lab = factor(ratioAB, levels = c(0.1, 0.5, 1, 2, 10),
                         labels = c("10:1", "2:1", "1:1", "1:2", "1:10"))
    ) %>%
  ggplot(aes(
    x = ID50,
    y = -log10(1 - neut),
    colour = ratio_lab
  )) +
  geom_line() +
  geom_line(data = titer_iip_pl, colour = "black", linetype = "dashed") +
  scale_x_log10(limits = c(0.5, 2500)) +
  scale_y_continuous(limits = c(NA, 6.2)) +
  facet_grid(cols = vars(interaction), rows = vars(IC50lab),
             switch = "x") +
  labs(x = "", colour = "conc. ratio", y = "IIP") +
  theme(legend.position = "top", strip.placement = "outside", 
        strip.background.x = element_blank())


```

